---
title: "Final Project"
author: "Paulo S"
date: "30 de noviembre de 2017"
output:
  slidy_presentation: default
  ioslides_presentation: default
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- This is an R Markdown presentation for presenting the unit converter. It appears on your left two boxes and an slider. You have to insert a numeric value in the boxes and press the buttom convert. You will see the conversion instead of NA values in the output panel. Aditionally, the slider helps you convert from celsius to farenheit degrees and you can see your conversion in the plot. Give a try¡

- Next, I am goin to show some examples.

## Input example

You'll see an input window like this...

```{r}
library(shiny)
ui<-library(shiny)
shinyUI(pageWithSidebar(
  headerPanel("Metric converter"),
  sidebarPanel(
    textInput(inputId="text1", label = "Farenheit"),
    textInput(inputId="text2", label = "Celsius"),
    sliderInput("Slider Temperature","Pick a value for celsius degrees",
                0,50,value = c(32)),
    submitButton("Convert!")
  ),
  mainPanel(
    p('Farenheit to Celsius'),
    textOutput('text1'),
    p('Celsius to Farenheit'),
    textOutput('text2'),
    p(''),
    plotOutput("plot1")
  )
))
server<-shinyServer(
  function(input, output) {
    output$text1 <- renderText({(as.numeric(input$text1)-32)*5/9})
    output$text2 <- renderText({(as.numeric(input$text2)*9/5)+32})
    output$plot1 <- renderPlot({
      pf <- c(input$`Slider Temperature`,(as.numeric(input$`Slider Temperature`)*9/5+32))
      curve(x*(9/5)+32, from=1, to=50,xlab="Celsius", ylab="Farenheit",
            main = "Farenheit - Celsius",lwd=2);grid()
      points(t(pf), pch=19,col="blue",cex=3)
      
    })
  }
)
shinyApp(ui, server)
```


## slider output

```{r,echo=F}
valueAssigned <- 10
pf <- c(valueAssigned,(as.numeric(valueAssigned)*9/5+32))
      curve(x*(9/5)+32, from=1, to=50,xlab="Celsius", ylab="Farenheit",
            main = "Farenheit - Celsius",lwd=2);grid()
      points(t(pf), pch=19,col="blue",cex=3)
```

## Thanks

```{r,echo=F}
curve(-x^2);grid()
```
